---
title: "Exercise Class Predictor"
author: "T Aristanto"
date: "Saturday, June 20, 2015"
output: html_document
---
## EXECUTIVE SUMMARY
This program is used to predict the class of exercise based on the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. For detail of the data plese see  http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 
Data was provided by  http://groupware.les.inf.puc-rio.br/har.
##DATA LOADING, SETUP AND PROCESS

Load Libraries
```{r load libraries,Echo=TRUE}
#For parallel processing
library(doParallel)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
#For Analysis and Machine Learning
library(AppliedPredictiveModeling)
library(caret)
```

We load the data from the given source. I choosed to download the data first before loading to R, thus the local loading folder. We cleaned the data in 3 ways:

- Remove the Not Available column
- Remove the first 7th column which contains unnecessary data
- Remove the Near Zero values

```{r Load , echo=TRUE, cache=TRUE}
RawTrain <- read.csv("C:/DataScience/Class8-Machine/pml-training.csv")
RawTest <- read.csv("C:/DataScience/Class8-Machine/pml-testing.csv")
exertrain<-RawTrain[,colSums(is.na(RawTrain)) == 0]
exertest <-RawTest[,colSums(is.na(RawTrain)) == 0]
exertrain <- exertrain[,-c(1:7)]
exertest <- exertest[,-c(1:7)]
nsvcol <- nearZeroVar(exertrain[,-ncol(exertrain)],saveMetrics=FALSE)
exertrain <- exertrain[-nsvcol]
exertest <- exertest[-nsvcol]
```

Check that the train data set has the same columns than that of test data set
```{r test column, echo=TRUE,cache=TRUE}
all.equal(colnames(exertest)[1:length(colnames(exertest))-1], colnames(exertrain)[1:length(colnames(exertrain))-1])
```


## ANALYSIS AND PREDICTION
We first create train and test data set out of the the training data set for Cross Validation. 
```{r Create train test data, echo=TRUE, cache=TRUE}
inTrain <- createDataPartition(exertrain$classe,p=0.75,list=FALSE)
training <- exertrain[inTrain,]
testing <- exertrain[-inTrain,]
```

We then use Random Forest Model to fit the data set
```{r Build Model, echo=TRUE, cache=TRUE}
set.seed(1404)
fitctrl <- trainControl(method="repeatedcv", number=3, repeats=3)
modelFit <- train(classe ~ ., data=training, method="rf", trControl=fitctrl)
modelFit$finalModel
```
 
To cross validate, we applied the model to the testing (from training set) prediction
```{r testing CV, echo=TRUE, cache=TRUE}
predtrain <- predict(modelFit, testing)
confusionMatrix(predtrain, testing$classe)
```
The model Shows a very high accuracy (over 99%), so this model is good.

## FINAL PREDICTION
We apply the model to the given test set
```{r final, echo=TRUE, cache=TRUE}
finalpred <- predict(modelFit, exertest)
finalpred

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(finalpred)
```

##REFERENCE
Please see: http://groupware.les.inf.puc-rio.br/har

Parallel processing code was provided by Russ Boucher from the class Forum. 
```